\textsf{Are frontier models superhuman chemists?}

### 2024-03-18 src/tex/ms.tex
update performance plot code
add references

Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\cite{White_2023}
This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\cite{jablonka202314, jablonka2024leveraging, xie2024fine} to optimize reactions\cite{ramos2023bayesian, kristiadi2024sober} or generate materials,\cite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\cite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\cite{bran2023chemcrow, Boiko_2023, darvish2024organa}
But even more than that, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a lot of untapped potential in \glspl{llm} for chemistry and materials science.
For instance, most insights in chemical research does not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
Most of these insights are in form of text in scientific publications. 
Thus, operating on such text might be our best way of unlocking and \emph{learning} from these insights.
This might ultimately lead to copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
Such a usage mode is especially interesting in the face of recent advanced in autonomous laboratories.\cite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology, that, for instance, can predict the toxicity of molecules (to be used in drug discovery) can also be used to predict the toxicity of molecules (to be used in chemical warfare).

@@

\begin{figure}
    \centering
    \includegraphics{figures/question_count_barplot.pdf}
    \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on a word-embedding of the question. 
    The figure shows that not all aspects of chemistry are equally represented in our corpus. The corpus currently focuses on safety-related aspects.}
    \label{fig:topic_barplot}
    \script{plot_statistics.py}
\end{figure}

@@

\begin{figure}
    \centering
    \caption{\textbf{Performance of models and humans on chembench.} The figure shows the percentage of questions that the models answered completely correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance. 
    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
    The metric we use here is harsh as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.}
    \label{fig:human_vs_models_bar}
\end{figure}

>>>>

Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\cite{White_2023}
This statement is motivated by a growing number of reports that use \glspl{llm} to properties of molecules or materials,\cite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking} to optimize reactions\cite{ramos2023bayesian, kristiadi2024sober} or generate materials,\cite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\cite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even build \enquote{autonomous} systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\cite{bran2023chemcrow, Boiko_2023, darvish2024organa}
But even more than that, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is a lot of untapped potential in \glspl{llm} for chemistry and materials science.
For instance, most insights in chemical research does not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
Most of these insights are in form of text in scientific publications. 
Thus, operating on such text might be our best way of unlocking and \emph{learning} from these insights.
This might ultimately lead to copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
Such a usage mode is especially interesting in the face of recent advanced in autonomous laboratories.\cite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}
However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\cite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology, that, for instance, can predict the toxicity of molecules (to be used in drug discovery) can also be used to predict the toxicity of molecules (to be used in chemical warfare).

@@

\begin{figure}
    \centering
    \includegraphics{figures/question_count_barplot.pdf}
    \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on a word-embedding of the question. 
    The figure shows that not all aspects of chemistry are equally represented in our corpus. The \chembench corpus, by design, currently focuses on safety-related aspects, which is also evident in \Cref{fig:question_diversity}. This figure represents the combined count of \glstext{mcq} and open-ended questions.}
    \label{fig:topic_barplot}
    \script{plot_statistics.py}
\end{figure}

@@

\begin{figure}
    \centering
    %\includegraphics{figures/human_vs_models_bar.pdf}
    \caption{\textbf{Performance of models and humans on chembench.} The figure shows the percentage of questions that the models answered completely correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance. 
    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
    The metric we use here is harsh as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.}
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}
